%
% ---- Bibliography ----
%


\begin{thebibliography}{}

\bibitem{datapic:xsd}
XML Schema 1.1, Datypic, http://www.datypic.com/sc/xsd11/ss.html

\bibitem{noauthor:ifc-guide}
The EXPRESS Definition Language for IFC Development, Stanford University documentation,
"http://web.stanford.edu/group/narratives/classes/08-09/CEE215/ReferenceLibrary/Industry Foundation Classes (IFC)/IFC General/The\_EXPRESS\_Definition\_Language\_for\_IFC\_Development.pdf"

\bibitem{w3c:ldp-best-practices}
Cody Burleson; Miguel Esteban Gutiérrez; Nandana Mihindukulasooriya. LDP Best Practices and Guidelines. W3C Working Draft. URL: http://www.w3.org/2012/ldp/hg/ldp-bp/ldp-bp.html

\bibitem{w3c:owl2-profiles}
World Wide Web Consortium. OWL 2 Web Ontology Language: Profiles (Second Edition), ed. Boris Motik et al. W3C Recommendation, 11 December 2012, http://www.w3.org/TR/2012/REC-owl2-profiles-20121211/. Latest version available at http://www.w3.org/TR/owl2-profiles/

\bibitem{w3c:xsd}
World Wide Web Consortium. W3C XML Schema Definition Language (XSD) 1.1 Part 2: Structures, ed. David Peterson et al. W3C Working Draft 3 December 2009. Available at: http://www.w3.org/TR/xmlschema11-2/

\bibitem{wiki:express}
EXPRESS (data modeling language), Wikipedia, http://en.wikipedia.org/wiki/EXPRESS\_(data\_modeling\_language)

\bibitem{w3c:owl2-syntax}
World Wide Web Consortium. OWL 2 Web Ontology Language 
Structural Specification and Functional-Style Syntax (Second Edition), ed. Boris Motik et al. W3C Recommendation, 11 December 2012, http://www.w3.org/TR/2012/REC-owl2-syntax-20121211/. Latest version available at http://www.w3.org/TR/owl2-syntax/

\bibitem{w3c:xmlschema11-2}
World Wide Web Consortium. W3C XML Schema Definition Language (XSD) 1.1 Part 2: Datatypes, ed. David Peterson et al. W3C Recommendation, 5 April 2012, http://www.w3.org/TR/2012/REC-xmlschema11-2-20120405#partial-implementation/, Section 5.4. Latest version available at http://www.w3.org/TR/xmlschema11-2/#partial-implementation

\bibitem{w3c:owl-guide}
World Wide Web Consortium. OWL Web Ontology Language Guide, ed. Michael K. Smith et al. W3C Recommendation, 10 February 2004, http://www.w3.org/TR/owl-guide/

\bibitem{w3c:rdf-schema}
World Wide Web Consortium. RDF Schema 1.1, ed. Dan Brickley et al. W3C Recommendation, 25 February 2014, http://www.w3.org/TR/rdf-schema

\end{thebibliography}
